A Large-scale Study of Polymorphism in R

### 2019-03-25 paper/main.tex
filled in some prelim numbers in corpus

@@

% https://www.r-pkg.org/downloaded
\begin{figure}[!th]{\footnotesize\begin{tabular}{@{}r||l|r|r|r@{}}\hline
\bf Package & \bf Description & \bf R LOC &\bf C LOC &\bf Scripts\\\hline
\tt Rcpp  & Seamless C++ integration & 2.1K & 4K & 12 \\
\tt rlang & Functions for 'Tidyverse' & 4.4K & 3.8K & ? \\
\tt glue  & Interpreted string literals & 0.2K & 0.2K & 8 \\
\tt tibble & Simple data frames & 1.3K & 0.3K & ? \\
\tt stringi &  String processing & 1.5K & 515K & ? \\
\tt ggplot2 & Data visualisations & 10.8K & 0 & 130 \\
\tt dplyr  &  Data manipulation & 3.8K & 4.7K & 78 \\
\tt pillar & Formatting for columns & 1.3K & 0 & ? \\
\tt R6 & Classes w. ref. semantics & 0.6K & 0 & ? \\
\tt stringr & String operations & 0.5K & 0 & ? \\
\end{tabular}}\caption{10 Most Downloaded Packages.}\label{most}
\end{figure}

@@

@@

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
function (object, fn = mean, par.name, comp.label, chain.no,  ...) 
{
    ...
    if (is.character(fn)) {
        if (fn == "MODE" | fn == "MAP") {
            do_MAP <- TRUE
        }
        else {
            do_MAP <- FALSE
            fn <- match.fun(fn) # looks up the function name
        }
    }
    ...
\end{lstlisting}}\caption{Example of char/closure argument (from {\tt BAMBI} package).}\label{fig:charclos}\end{figure}

@@

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
nret.translator <- function(dat, spss="out", ...){
  ...
  if(identical(spss, "out")) {
    if(!is.vector(dat)) {
        stop(simpleError("...'dat' must be a vector!"))
      }
    ...
  } else {
    ...
    # here, dat is data.frame
    items.idx <- items.idx[order(names(dat[, items.idx]))]
  }
\end{lstlisting}}\caption{Example of a data.frame/(vector/double) argument (from {\tt klausR} package).}\label{fig:dfdbl}\end{figure}

@@

In Table~\ref{tab:toppolyFDTS}, we see the prominence of {\tt double} and {\tt integer} polymorphism.
In many dynamic languages, primitive double and integer types can easily stand-in for each other as coercion between them is easy.
In R, the distinction is mainly just relevant to the implementation, as integers and doubles are stored and dealt with differently in the runtime but programmers only see the distinction when printing.
We address this by creating a new type, {\it real}, encompassing both.
The expected coercion is evident throughout R.
Consider the snippet in Figure~\ref{fig:realex} for details.
\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
> 5L + 1L # => 6L, an integer
> 5L + 1.2 # => 6.2, a double (5L coerced to 5.0)
> c(10, 20, 30)[1.2] # => 10, 1.2 coerced to 1L
\end{lstlisting}}\caption{Example of {\it real} type usage.}\label{fig:realex}\end{figure}

>>>>

\AT{TODO Change "recorded" to "observed" in recordsbypkg}

@@

% https://www.r-pkg.org/downloaded
\begin{figure}[!th]{\footnotesize\begin{tabular}{@{}r||l|r|r|r|r|r@{}}\hline
\bf Package & \bf Description & \bf R LOC &\bf C LOC &\bf Scripts & \bf Calls Observed & \bf Calls Recorded \\
\hline
\tt Rcpp  & Seamless C++ integration & 2.1K & 4K & 25 & 51K & 282 \\
\tt rlang & Functions for 'Tidyverse' & 4.4K & 3.8K & 122 & 2,240K & 5,171 \\
\tt glue  & Interpreted string literals & 0.2K & 0.2K & 8 & 2K & 120 \\
\tt tibble & Simple data frames & 1.3K & 0.3K & 16 & 839K & 4,148 \\
\tt stringi &  String processing & 1.5K & 515K & 64 & 692K & 666 \\
\tt ggplot2 & Data visualisations & 10.8K & 0 & 130 & 110K & 2,636\\
\tt dplyr  &  Data manipulation & 3.8K & 4.7K & 78 & 157K & 1,967 \\
\tt pillar & Formatting for columns & 1.3K & 0 & 13 & 549K & 1,273 \\
\tt R6 & Classes w. ref. semantics & 0.6K & 0 & 2 & 1K & 226 \\
\tt stringr & String operations & 0.5K & 0 & 32 & 1,551K & 371 \\
\end{tabular}}\caption{10 Most Downloaded Packages.}\label{most}
\end{figure}

@@

\AT{Putting examples of polymorphism in this section, we can talk about them and where they should go.}

First, Figure~\ref{fig:realex} shows an example of how integers and doubles are very compatible.
\AT{TODO add matrix (dimensions) to the example.}

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
> 5L + 1L # => 6L, an integer
> 5L + 1.2 # => 6.2, a double (5L coerced to 5.0)
> c(10, 20, 30)[1.2] # => 10, 1.2 coerced to 1L
\end{lstlisting}}\caption{Example of {\it real} type usage.}\label{fig:realex}\end{figure}

@@

Figure~\ref{fig:charclos} shows an example of a function with character and function polymorphism.
\AT{Bluh.}
At a high level, this function calculates point estimates for an {\tt angmcmc} object (specific to some packages).
The caller specifies {\tt fn}, either a function or a function name which will be evaluated on the object samples to estimate parameters.
The call to lookup the function name is on line 10.

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
pointest <- function (object, fn = mean, par.name, comp.label, chain.no,  ...) {
    ...
    if (is.character(fn)) {
        if (fn == "MODE" | fn == "MAP") {
            do_MAP <- TRUE
        }
        else {
            do_MAP <- FALSE
            fn <- match.fun(fn) # looks up the function name
        }
    }
    ...
\end{lstlisting}}\caption{Example of char/closure argument (from {\tt BAMBI} package).}\label{fig:charclos}\end{figure}

@@

\begin{figure}[!hb]{\small\begin{lstlisting}[style=R]
nret.translator <- function(dat, spss="out", ...) {
  ...
  if(identical(spss, "out")) {
    if(!is.vector(dat)) {
        stop(simpleError("...'dat' must be a vector!"))
      }
    ...
  } else {
    ...
    # here, dat is data.frame
    items.idx <- items.idx[order(names(dat[, items.idx]))]
  }
\end{lstlisting}}\caption{Example of a data.frame/(vector/double) argument (from {\tt klausR} package).}\label{fig:dfdbl}\end{figure}

@@

In Table~\ref{tab:toppolyFDTS}, we see the prominence of {\tt double} and {\tt integer} polymorphism.
In many dynamic languages, primitive double and integer types can easily stand-in for each other as coercion between them is easy.
In R, the distinction is mainly just relevant to the implementation, as integers and doubles are stored and dealt with differently in the runtime but programmers only see the distinction when printing.
We address this by creating a new type, {\it real}, encompassing both.
The expected coercion is evident throughout R.
Refer to the snippet in Figure~\ref{fig:realex} for details.
